# Title

```{r}
require(knitr)
require(MASS)
require(data.table)
require(ggplot2)
```


## Definitions

_Gaussian process_ is a collection of random variables, any
finite number of which have a joint Gaussian distribution.

A Gaussian process is completely specified by its mean function and
co- variance function. We define mean function \(m(x)\) and the
covariance function \(k(x,x')\) of a real process \(f(x)\) as
_covariance_ and _mean_ function
\[m(x) = \mathcal{E}[f(x)],\]
\[ k(x,x') = \mathcal{E}[(f(x) − m(x))(f(x') − m(x'))],\]
and will write the Gaussian process as
\( f(x) \sim GP(m(x), k(x, x'))\).

## Squared exponential

\[ cov(f(x_p), f (x_q)) = k(x_p , x_q) =
    \exp\Big(−\frac{1}{2} |x_p − x_q |\Big)^2.\]

```{r}
x <- seq(-5,5,0.5)

sim.number <- 5


squared.exp <- function(x,y,l=1) {
  exp(-0.5* (norm(x-y, type="2")/l)^2)
}

Sigma <- function(x, l=1) {
  n <- length(x)
  S <- matrix(rep(0, n*n), nrow=n)
  for (i in 1:n) {
    for (j in 1:n) {
      S[i,j] <- squared.exp(x[i],x[j],l)
    }
  }
  S
}

S <- Sigma(x)
sims <- data.table(x=x)
gens <- mvrnorm(sim.number, rep(0, length(x)), S)
sims <- cbind(sims, t(gens))
ggplot(sims,aes(x)) + geom_line(aes(y=V1)) +
    geom_line(aes(y=V2)) +
        geom_line(aes(y=V3)) 
```


